---
title: "lecture9"
author: "Anastasia Pimentel"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##K-means clustering 

Generating data for our cluster:
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

Now we use k-means:
```{r}
k <- kmeans(x,centers=2, nstart=20)
k
```

Q. How many points are in each cluster?
  There are 30 points each.
Q. What ‘component’ of your result object details
      - cluster size? - Use "size."
      - cluster assignment/membership? Use "cluster."
      - cluster center? Use "centers."



Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
```{r}

plot(x, col=k$cluster)
points(k$centers, col="blue")
```

##Hierarchical clustering
Now we'll practice hierarchical clustering using the hclust function. 


```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)


plot(hc)
abline(h=6, col="red", lty=2)


cutree(hc, h=6)
```


What about unlcear data?


```{r}
# Step 1. Generate some example data for clustering
x2 <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x2) <- c("x2", "y")
# Step 2. Plot the data without clustering
plot(x2)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x2, col=col)
```

Now to find cluster membership and tabulate them.
```{r}
hc2 <- hclust(dist(x2))
plot(hc2)
abline(h=2, col="red", lty=2)
ct <- cutree(hc2, h=2)
table(ct)

#Now plot with colors for the clusters.

plot(x2, col=ct)
```


##PCA
Now we'll practice using PCA on lots of data.
```{r}
x <- read.csv("UK_foods.csv", row.names=1)
x

```
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
pairs(x, col=rainbow(10), pch=16)

```

Principal Component Analysis (PCA) with 
```{r}

pca <- prcomp( t(x) )
summary(pca)
#The summary isn't very useful, so lets dig deeper.
#Checking the attribute of pca. "x" is the new variables that the PCA funtion creates. 
attributes(pca)

```
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=c("black", "red", "blue", "darkgreen"))
```

So N. Ireland is different. But different in what way?
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

